Peer Assessment 1 of Reproducible Research course on Coursera
by M. Sabelnikov


## Loading and preprocessing the data
Loading packages adn setting time settings
```{r, results='hide',message=FALSE}
library(dplyr)
library(data.table)
library(ggplot2)
Sys.setlocale("LC_TIME", "USA")
```
Reading data from CSV in working directory:
```{r}
data<-read.csv("activity.csv")
```

Checking the structure of the data:

```{r}
str(data)
```
## What is mean total number of steps taken per day?

Calculating mean, sum and median by day
```{r}
mean_steps<-data %>% group_by(date) %>% summarize(steps=mean(steps,na.rm=T))
sum_steps<-data %>% group_by(date) %>% summarize(steps=sum(steps,na.rm=T))
median_steps<-data %>% group_by(date) %>% summarize(steps=median(steps,na.rm=T))
```
Make a histogram of the total number of steps taken each day
```{r}
hist(sum_steps$steps,main="Sum of steps by day (Without NA values)",xlab="Steps")
```

Mean and meadian steps by day table
```{r}
mean_steps<-data.table(mean_steps)
mean_steps<-cbind(mean_steps,median_steps$steps)
setnames(mean_steps,c("date","mean","median"))
mean_steps
```

Mean number of steps is **`r mean(data$steps,na.rm=T)`**  
Median number of steps is **`r median(data$steps,na.rm=T)`**

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Cleaning data
```{r}
data_clean1<-complete.cases(data$steps)
data_clean<-data[data_clean1,]
```

Calculating summary table and plotting
```{r}
interval<-group_by(data_clean,interval)
int<-summarize(interval,steps=mean(steps))
plot(int$interval,int$steps,type="l",xlab="Interval",ylab="Steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
m<-int[int$steps==max(int$steps),1]
```

The interval with maximum number of steps is interval # **`r m`**

## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
nas<-table(data_clean1)
```
The total number of missing values in the dataset is **`r as.integer(nas[1])`**

Filling in all of the missing values in the dataset

```{r}
data_wo_nas <-merge(data,int,by="interval")

id<-1:17568
for (i in id){
        if (is.na(data_wo_nas[i,2])){
                data_wo_nas[i,2]<-data_wo_nas[i,4]}
        else {}
        
}

```

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
data_new<-select(data_wo_nas,steps=steps.x,interval,date)
```

Make a histogram of the total number of steps taken each day
```{r}
sum_steps_new<-data_new %>% group_by(date) %>% summarize(steps=sum(steps,na.rm=T))
hist(sum_steps_new$steps,main="Sum of steps by day (with NAs reaplaces by mean)",xlab="Steps")
```
  
**Here is initial histogram**
```{r}
hist(sum_steps$steps,main="Sum of steps by day (Without NA values)",xlab="Steps")
```

Mean number of steps in data with NAs replaced is **`r mean(data_new$steps,na.rm=T)`**  
Median number of steps in data with NAs replaced is **`r median(data_new$steps,na.rm=T)`**  

As we can see total mean didn't change from the estimation of the 1st part of assignment.

## Are there differences in activity patterns between weekdays and weekends?

Adding column with weekdays
```{r}
data_new$weekday<-weekdays(as.Date(data_new$date))
data_new$weekday<-ifelse(data_new$weekday %in% c("Saturday", "Sunday"),"weekend", "weekday")
```

Calculate averages and plot.
```{r}
data_new2<-data_new %>% group_by(interval,weekday) %>% summarize(steps=mean(steps,na.rm=T))
ggplot(data_new2, aes(x=interval, y=steps, fill=weekday)) +
        geom_line(stat="identity") + 
        facet_grid(weekday  ~ ., scales="free") +
        ylab("Steps") + 
        xlab("Intervals") +
        ggtitle(expression("Average number of steps taken"))

```
